### 2014-04-09 binary/LaTeX/RareBinary.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amssymb, amsthm, amsmath, amsfonts}
\usepackage{doublespace}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}

\linespread{1.0}

\newcommand{\btheta}{ \mbox{\boldmath $\theta$}}
\newcommand{\bmu}{ \mbox{\boldmath $\mu$}}
\newcommand{\balpha}{ \mbox{\boldmath $\alpha$}}
\newcommand{\bbeta}{ \mbox{\boldmath $\beta$}}
\newcommand{\bdelta}{ \mbox{\boldmath $\delta$}}
\newcommand{\blambda}{ \mbox{\boldmath $\lambda$}}
\newcommand{\bgamma}{ \mbox{\boldmath $\gamma$}}
\newcommand{\brho}{ \mbox{\boldmath $\rho$}}
\newcommand{\bpsi}{ \mbox{\boldmath $\psi$}}
\newcommand{\bepsilon}{ \mbox{\boldmath $\epsilon$}}
\newcommand{\bomega}{ \mbox{\boldmath $\omega$}}
\newcommand{\bDelta}{ \mbox{\boldmath $\Delta$}}
\newcommand{\bSigma}{ \mbox{\boldmath $\Sigma$}}
\newcommand{\bA}{ \mbox{\bf A}}
\newcommand{\bP}{ \mbox{\bf P}}
\newcommand{\bx}{ \mbox{\bf x}}
\newcommand{\bX}{ \mbox{\bf X}}
\newcommand{\bB}{ \mbox{\bf B}}
\newcommand{\bZ}{ \mbox{\bf Z}}
\newcommand{\by}{ \mbox{\bf y}}
\newcommand{\bY}{ \mbox{\bf Y}}
\newcommand{\bz}{ \mbox{\bf z}}
\newcommand{\bh}{ \mbox{\bf h}}
\newcommand{\br}{ \mbox{\bf r}}
\newcommand{\bt}{ \mbox{\bf t}}
\newcommand{\bs}{ \mbox{\bf s}}
\newcommand{\bb}{ \mbox{\bf b}}
\newcommand{\bL}{ \mbox{\bf L}}
\newcommand{\bu}{ \mbox{\bf u}}
\newcommand{\bv}{ \mbox{\bf v}}
\newcommand{\bV}{ \mbox{\bf V}}
\newcommand{\bG}{ \mbox{\bf G}}
\newcommand{\bH}{ \mbox{\bf H}}
\newcommand{\bw}{ \mbox{\bf w}}
\newcommand{\bo}{ \mbox{\bf o}}
\newcommand{\bfe}{ \mbox{\bf e}}
\newcommand{\iid}{\stackrel{iid}{\sim}}
\newcommand{\indep}{\stackrel{indep}{\sim}}
\newcommand{\calR}{{\cal R}}
\newcommand{\calG}{{\cal G}}
\newcommand{\calD}{{\cal D}}
\newcommand{\calS}{{\cal S}}
\newcommand{\calB}{{\cal B}}
\newcommand{\calA}{{\cal A}}
\newcommand{\calT}{{\cal T}}
\newcommand{\calO}{{\cal O}}
\newcommand{\argmax}{{\mathop{\rm arg\, max}}}
\newcommand{\argmin}{{\mathop{\rm arg\, min}}}
\newcommand{\Frechet}{\mbox{Fr$\acute{\mbox{e}}$chet}}
\newcommand{\Matern}{ \mbox{Mat$\acute{\mbox{e}}$rn}}

\newcommand{\beq}{ \begin{equation}}
\newcommand{\eeq}{ \end{equation}}
\newcommand{\beqn}{ \begin{eqnarray}}
\newcommand{\eeqn}{ \end{eqnarray}}

\begin{document}\linenumbers

\begin{center}
{\Large {\bf A spatial model for rare binary events}}\\
\today
\end{center}

\section{Introduction}\label{s:intro}

\section{New Model?}\label{s:model}

Let $Y_i\in\{0,1\}$ be the binary response at spatial location $\bs_i\in\calD$, and $\bX_i$ be the associated $p$-vector of covariates with first element equal to one for the intercept.  We relate the covariates with the response using the link function $g$ so that P$(Y_i=1) = p_i= g(\bX_i\bbeta)$, where $\bbeta$ is the $p$-vector of regression coefficients.  For example, \cite{Wang-2010} propose the GEV link function
$p_i=1-\exp\left[\left(1-\xi\bX_i\bbeta\right)^{-1/\xi}\right]$ for rare binary data. We will also consider logit and probit links.

-- Not quite sure why the article uses this. I think we should use
\begin{align}
	p_i = 1 - \exp \left[ - \left(1 + \xi \bX_i \bbeta \right)^{-1/\xi} \right] 
\end{align}

We propose a copula \citep{nelsen-1999} to account for spatial dependence while preserving the marginal event probabilities. Let $Y_i = I(Z_i>z_i)$, where $Z_i$ is a continuous latent variable and $z_i$ is the appropriate threshold so that $P(Y_i=1)=p_i$.  The latent $Z_i$ is modeled using spatial extreme value analysis methods to capture dependence between rare events.  We assume $Z$ follows the max-stable spatial process of \cite{reich-2012}.  Under this model, the marginal distribution of each $Z_i$ is GEV(1,1,1) with P$(Z_i>c) = 1-\exp(-1/c)$.  Therefore, we must set $z_i=-1/\log(1-p_i)$ so that $P(Y_i=1)=p_i$.

Spatial dependence is determined by the joint distribution of $\bZ = (Z_1,\ldots,Z_n)$, 
\beq\label{jointCDF}
 G(\bz) =  \mbox{P}[Z_1<z_1,\ldots,Z_n<z_n] = \exp\left\{-\sum_{l=1}^L\left[\sum_{i=1}^n\left(\frac{w_{l}(\bs_i)}{z_i}\right)^{1/\alpha}\right]^{\alpha}\right\},
\eeq
where $\bz = (z_1,\ldots,z_n)$. This is a special case of the multivariate GEV distribution with asymmetric Laplace dependence function \citep{Tawn-1990}.  The parameter $\alpha\in(0,1)$ determines the strength of dependence, with $\alpha$ near zero giving strong dependence and $\alpha=1$ giving joint independence. The weights $w_{li}>0$ determine the spatial dependence structure, and are discussed in detail in Section \ref{s:spatial}.  Many weight functions are possible, but the weights must be constrained so that $\sum_{l=1}^L w_{l}(\bs_i)=1$ for all $i=1,\ldots,n$ to preserve the marginal GEV distribution.

\section{Spatial dependence}\label{s:spatial}

The weights $w_{l}(\bs_i)$ in (\ref{jointCDF}) should vary smoothly across space to induce spatial dependence.  For example, \cite{reich-2012} take the weights to be scaled Gaussian kernels with knots $\bv_l$, that is
\beq\label{w}
   w_{l}(\bs_i) = \frac{\exp\left[-0.5\left(||\bs_i-\bv_l||/\rho\right)^2\right]}
                 {\sum_{j=1}^L\exp\left[-0.5\left(||\bs_i-\bv_j||/\rho\right)^2\right]}.
\eeq
To kernel bandwidth $\rho>0$ determines the spatial range of the dependence, with large $\rho$ giving long-range dependence and vice versa.

Then in a bivariate setting, the probability of observing a joint exceedances as a function of $\alpha$ is

\begin{align} \label{biv}
  \text{P}(Y_i = 1, & Y_j = 1) = 1 - \exp\left\{ - \frac{ 1 }{ z_i } \right\} - \exp \left\{ - \frac{ 1 }{z_j} \right\} + \exp \left\{ - \sum_{ l = 1 }^{ L } \left[ \left( \frac{ w_{ l }(\bs_i) }{ z_i } \right)^{1/\alpha} + \left( \frac{ w_{l }(\bs_i)}{z_j} \right)^{1/\alpha} \right]^{\alpha} \right\} \nonumber \\
  		&= p_i + p_j - \left( 1 - \exp \left \{ - \sum_{ l = 1 }^{ L } \left( \left[ -\log(1-p_i) w_{l}(\bs_i) \right]^{ 1/\alpha } + \left[ -\log(1 - p_j) w_{ l}(\bs_j) \right]^{ 1/\alpha } \right)^{\alpha} \right \} \right).
\end{align}

To describe the tail dependence, we use the $\chi$ statistic of \citet{Coles-1999}.
Assume that $Y_i$ and $Y_j$ have the same marginal distributions, then $p_i = p_j = p$ for all $i, j$.
As shown in Appendix A.2,

\begin{align}
\chi = 2 -  \vartheta(\bs_i, \bs_j).
\end{align}
where $\vartheta (\bs_i, \bs_j) = \sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i)^{ 1/\alpha } +  w_{ l}(\bs_j)^{ 1/\alpha } \right]^\alpha$ is the pairwise extremal coefficient given by \citet{reich-2012}.
In the case of complete dependence, $\chi = 1$, and in the case of complete independence, $\chi = 0$. 
This is relatively easy to show for $\alpha = 1$, but I don't know of a way to prove $\lim_{\alpha \rightarrow 0} \chi = 1$. Any thoughts?

\section{Computation}\label{s:comp}

As shown in Appendix A.1, the joint probability mass function of $\bY=(Y_1,\ldots,Y_n)$ has a convenient form when the number of events is small.  Let $K=\sum_{i=1}^nY_i$ be the number of events, and assume without loss of generality the data are ordered so that the $Y_1=\ldots=Y_K=1$.  Then
\beq\label{pmf}
   P(Y_1=y_1,\ldots,Y_n=y_n) =  \left\{
                               \begin{array}{ll}
                                 G(\bz) & K=0 \\
                                 G(\bz_{(1)})-G(\bz) & K=1 \\
                                 G(\bz_{(12)})-G(\bz_{(1)})-G(\bz_{(2)})+G(\bz) & K=2
                               \end{array}
                             \right.
\eeq
where $G(\bz_{(1)}) = P(Z_2<z_2,\ldots,Z_n<z_n)$, $G(\bz_{(2)}) = P(Z_1<z_1,Z_3<z_3,\ldots,Z_n<z_n)$, and $G(\bz_{(12)}) = P(Z_3<z_3,\ldots,Z_n<z_n)$.  Similar expressions can be derived for all $K$, but become cumbersome for large $K$.  Therefore, for small $K$ we can evaluate the likelihood directly.  Most days in our dataset have $K<4$, so we use this expression for those days.  However for days with many events, we must use the latent variable scheme described below (unless you can think of a better way!).
I think it should be more computationally efficient to use (\ref{pmf}) for any $K$. At most, we have to calculate the $\left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{1/\alpha}$, for all $i, l$. In the random effects model, the expression for the joint density conditional on $\theta$ is 
\beq
	P(Y_1=y_1,\ldots,Y_n=y_n) = \prod_{ i = 1 }^{ n } \left[ \exp \left\{ \sum_{ l = 1 }^{L} A_l \left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{ 1/\alpha} \right\} \right]^{ 1 - Y_i } \left[ 1 - \exp \left\{ \sum_{ l = 1 }^{L} A_l \left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{ 1/\alpha} \right\} \right]^{ Y_i }.
\eeq
So we still need to compute $\left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{1/\alpha}$, but we also need to do the sampling for all the $A_l$ terms as well.

\section{Simulation study}\label{s:sim}

\section{Data analysis}\label{s:analysis}

\section{Conclusions}\label{s:con}

\section*{Acknowledgments}

\section*{Appendix A.1: Derivation of the likelihood}
We use the hierarchical max-stable spatial model given by \citet{reich-2012}. If at each margin, $Z_i \sim $ GEV$(1,1,1)$, then $Z_i | \theta_i \indep $ GEV$(\theta, \alpha \theta, \alpha)$. As defined in section \ref{s:comp}, we reorder the data such that $Y_1=\ldots=Y_K=1$, and $Y_{K+1} = \ldots = Y_n = 0$. Then the joint likelihood conditional on the random effect $\theta$ is

\begin{align} \label{joint_cond}
	P(Y_1=y_1,\ldots,Y_n=y_n) &= \prod_{ i \le K } \left\{ 1 - \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \right \} \prod_{ i > K } \exp \left[ -\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \nonumber \\[0.5em]
		&= \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] - \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ i = 1}^{K} \exp\left[ -\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \nonumber\\
		&  + \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ 1 < i < j \le K } \left\{ \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} - \left( \frac{ \theta_j }{ z_j } \right)^{ 1/\alpha } \right] \right \} \nonumber \\[0.5em]
		& + \cdots + (-1)^K \exp\left[ - \sum_{ i = 1 }^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right]
\end{align}

Finally marginalizing over the random effect, we obtain

\begin{align} \label{joint}
    P(Y_1=y_1,\ldots,Y_n=y_n) &=\int G(\bz | \bA) p( \bA | \alpha) d\bA. \nonumber\\[0.5em]
			&= \int \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] - \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ i = 1}^{K} \exp\left[ -\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \nonumber\\
		&  + \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ 1 < i < j \le K } \left\{ \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} - \left( \frac{ \theta_j }{ z_j } \right)^{ 1/\alpha } \right] \right \} \nonumber \\[0.5em]
		& + \cdots + (-1)^K \exp\left[ - \sum_{ i = 1 }^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] p( \bA | \alpha) d\bA.
\end{align}

Consider the first term in the summation,

\begin{align}
	\int \exp \left\{ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right\} p( \bA | \alpha) d\bA &= \int \exp \left\{ - \sum_{ i = K + 1 }^n \left( \frac{ \left[ \sum_{ l = 1 }^L  A_l w_{l}(\bs_i)^{1/\alpha} \right)^\alpha }{ z_i} \right]^{ 1/\alpha } \right \} p( \bA | \alpha) d\bA \nonumber \\[0.5em]
	 &= \int \exp \left\{ -\sum_{ i = K + 1}^n \sum_{ l = 1}^L A_l \left( \frac{ w_l (\bs_i) }{ z_i } \right)^{1/\alpha} \right \} p( \bA | \alpha) d\bA \nonumber \\[0.5em]
	 &=\exp\left\{-\sum_{ l = 1}^L \left[ \sum_{ i = K + 1 }^n \left( \frac{ w_l(\bs_i)}{ z_i} \right)^{1/\alpha} \right]^\alpha \right\}.
\end{align}

The remaining terms in equation (\ref{joint}) are straightforward to obtain, and after integrating out the random effect, the joint density is the density given in (\ref{pmf}).

\section*{Appendix A.2: Derivation of the $\chi$ statistic}
\begin{align} \label{chi}
  \chi &= \lim_{p \rightarrow 0 }\text{P}(Y_i = 1|Y_j = 1)\nonumber \\
   &= \lim_{p \rightarrow \infty }\frac{ p + p - \left( 1 - \exp \left \{ - \sum_{ l = 1 }^{ L } \left[ \left( -\log(1-p) w_{l}(\bs_i) \right)^{ 1/\alpha } + \left( -\log(1 - p) w_{ l}(\bs_j) \right)^{ 1/\alpha } \right]^{\alpha} \right \} \right) }{ p } \nonumber \\[0.5em]
  &= \lim_{p \rightarrow 0 } \frac{ 2p - \left( 1 - \exp \left \{ \log(1-p) \sum_{ l = 1 }^{ L } \left[  w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l}(\bs_j) ^{ 1/\alpha } \right]^{\alpha} \right \} \right) }{ p } \nonumber \\[0.5em]
  &= \lim_{p \rightarrow 0 } \frac{ 2p - \left( 1 - (1-p)^{ \sum_{ l = 1 }^{ L } \left[ \left( w_{l}(\bs_i) \right)^{ 1/\alpha } + \left( w_{ l}(\bs_j) \right)^{ 1/\alpha } \right]^\alpha } \right) }{ p } \nonumber \\[0.5em]
  &=\lim_{p \rightarrow 0 } 2 -\sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l} (\bs_j)^{ 1/\alpha } \right]^\alpha ( 1 - p)^{ -1 + \sum_{ l = 1 }^{ L } \left[  w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l}(\bs_j)^{ 1/\alpha } \right]^\alpha } \nonumber \\[0.5em]
  &= 2 -  \sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i)^{ 1/\alpha } +  w_{ l}(\bs_j) ^{ 1/\alpha } \right]^\alpha.
\end{align}

\begin{singlespace}
\bibliographystyle{rss}
\bibliography{spatialbinary}
\end{singlespace}

\end{document}