### 2014-09-08 pot/LaTeX/proof_dist.tex
minor changes to manuscript.

Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$. Let $\calD = B_{2h}(\bm)$ which is a ball of radius $h > 0$ centered at $\bm$, the midpoint between $\bs_1$ and $\bs_2$, and $h = ||\bs_1 - \bs_2||$. Let $w_1, \ldots, w_K \sim \text{Unif}(\calD)$ be a set of random knots uniformly distributed over $\calD$. Define $g(\bs_i) = \argmin_j || \bs_i - w_j ||$ to be the partition in which $\bs_i$ resides. Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2)]$ be the probability that both sites are in the same partition. Then $\pi(h)$ is the probability that $w_j$ is the only knot in $(B_a(\bs_1) \cup B_b(\bs_2) \cap \calD)$ where $a = ||\bs_1 - w_j||$ and $b = ||\bs_2 - w_j||$.