### 2014-12-09 pot/LaTeX/proof_dist.tex
Revert d132b13..1d41aaf
This rolls back to commit d132b1363ba8f49558713cb8aa9a840bfeb35ba0.

Consider two spatial locations, $\bs_1, \bs_2 \in \calR^2$, and any two knots, $w_1, w_2 \in \calR^2$.
Let $\overline{S}$ be the line segment connecting $\bs_1$ and $\bs_2$.
Let $h = ||\bs_1 - \bs_2||$ be the distance between sites $\bs_1$ and $\bs_2$.
Let $\overline{W}$ be the perpendicular bisector of the line connecting $w_1$ and $w_2$.
Then $\overline{W}$ defines a partition dividing $\calR^2$ into two half-planes.
Let $\lim_{h \rightarrow \infty}$ be a limit in the sense that both $\bs_1$ and $\bs_2$ are moving at some rate away from their midpoint.
Provided that $\overline{W}$ and $\overline{S}$ are not parallel to one another, we are guaranteed that as $h \rightarrow \infty$, $\bs_1$ and $\bs_2$ will be in different partitions.