### 2014-12-09 pot/LaTeX/proof_dist.tex
Revert 1d41aaf..e1a4cbd
This rolls back to commit 1d41aafe4e88de9d164d76b14831c62d29f642a2.

% Consider two spatial locations, $\bs_1, \bs_2 \in \calR^2$, and any two knots, $w_1, w_2 \in \calR^2$.
% Let $\overline{S}$ be the line segment connecting $\bs_1$ and $\bs_2$.
% Let $h = ||\bs_1 - \bs_2||$ be the distance between sites $\bs_1$ and $\bs_2$.
% Let $\overline{W}$ be the perpendicular bisector of the line connecting $w_1$ and $w_2$.
% Then $\overline{W}$ defines a partition dividing $\calR^2$ into two half-planes.
% Let $\lim_{h \rightarrow \infty}$ be a limit in the sense that both $\bs_1$ and $\bs_2$ are moving at some rate away from their midpoint.
% Provided that $\overline{W}$ and $\overline{S}$ are not parallel to one another, we are guaranteed that as $h \rightarrow \infty$, $\bs_1$ and $\bs_2$ will be in different partitions.
Consider two spatial locations $\bs_1, \bs_2 \in \calR^2$, and consider spatial knots that follow a Poisson process with intensity $\lambda$.
Let $h = ||\bs_1 - \bs_2||$, and let $\pi(h)$ be the probability that two sites are in the same partition.
To guarantee that $\bs_1$ and $\bs_2$ are indifferent partitions, we need at least two knots to be in between $\bs_1$ and $\bs_2$.
Let $B_{\bs}(\frac{h}{2})$ be a ball that is of radius $\frac{h}{2}$ that has $\bs_1$ and $\bs_2$ on its circumference, and consider $\overline{S}$ to be the line connecting $\bs_1$ and $\bs_2$ that divides $B_{\bs}(\frac{h}{2})$ into two sizes of equal area.