### 2015-02-17 LaTeX/MVN.tex
change to file organization

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{1.0}
\input{mycommands.sty}

\begin{document}\linenumbers

\begin{center}
{\Large {\bf A new spatial model for points above a threshold}}\\
\today
\end{center}

MVN results:
The distribution of $\bx_1$ conditional on $\bx_2 = a$ is $N( \mu^*, \Sigma^*)$ where
\begin{align*}
  \mu^* &= \mu_1 + \Sigma_{12} \Sigma_{22}^{-1} (a - \mu_2)\\
  \Sigma^* &= \Sigma_{11} - \Sigma_{12} \Sigma_{22}^{-1} \Sigma_{21}
\end{align*}

\end{document}