### 2015-05-31 markdown/exploring_model_fit.Rmd

---
title: "Exploring Model Fits"
author: "Sam Morris"
date: "19 May 2015"
output: pdf_document
---

```{r, echo=FALSE, include=FALSE}
load("./code/analysis/simstudy/simdata.RData")
library(evd)  # to load in chiplot
```

The goal of this document is to try to explore different ways to think about dependence in the data.
In talking with Emeric and Dan, it's not clear to me that we have the best way to assess model fit.
Brier scores are nice in that they give a single number, and similar to other decision making criterion, you can decide on the best model simply by looking at which model has the lowest Brier score.
Furthermore, the Brier score is the proper scoring rule that can be used to assess how well you do at predicting the occurence of an event.

One thing I've been thinking about is trying to figure out how well we do at recreating the time series for a specific site. 
One of the main goals of our model is to predict at unobserved locations.
Therefore, it would be desirable to know how well we do at estimation of the time series at an unobserved location.
One plot that might be interesting is the posterior distribution at a particular site for all of the time points.
If we overlay the observed values, you could see how we do marginally for 3 or 4 sites.
You could also find some way to differentiate the values that exceed the 95th quantile (for the ozone data, the top 2 days; simulated data, the top 3 days).